---
title: "Plant functional group composition and vegetation structure at CGREC"
author: DAM
date: '`r Sys.Date()`'
output: 
  html_document:
    theme: united
    highlight: tango
    toc: TRUE
---

```{r setup, echo=FALSE, warning=FALSE, message = FALSE, results='hide'}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=FALSE, eval=TRUE) 
# fig.path='../figures/', dev=c('png', 'pdf')
# Packages 
  pacman::p_load(tidyverse, vegan, RVAideMemoire)
  pacman::p_load_gh('devanmcg/wesanderson')
  source("https://raw.githubusercontent.com/devanmcg/IntroRangeR/master/11_IntroMultivariate/ordinationsGGplot.R")
```


```{r data_loading}
# Load and arrange data 
veg_all <-
  read_csv('../data/CGRECVegCompCameron.csv') %>% 
    mutate(block = substr(transect, 1,1), 
           pasture = substr(transect, 2,3), 
           patch = substr(transect, 4,5), 
           subpatch = substr(transect, 6,6), 
           transect = paste(block, pasture, patch, subpatch, sep = "."), 
           pasture = paste(block, pasture, sep = "_")) %>%
    select(block, pasture, patch, subpatch, transect, 
           TRT, year, TSF, BurnSeason, 
           Bare:LitterDepth, 
           POPR:IntWoody) %>%
    mutate( TRT = factor(TRT, levels = c("PBG40", "PBG20", "SLG", "WAG")), 
            TRT = recode(TRT, 
                        "PBG40" = "1 burns/yr", 
                        "PBG20" = '2 burn/yr',
                        "SLG" = "Continuous", 
                        "WAG" = "Rotational") )
# Alternatively, with relative dominance
veg_ra <- 
  veg_all %>%
  rowwise() %>%
  mutate(TotalCover = sum(c_across(POPR:IntWoody)), 
         across(POPR:IntWoody, ~((./TotalCover)*100))) %>%
  ungroup() %>%
  filter(TotalCover != 0.0) %>%
  select(-TotalCover) 

# Patch contrast 
  load('../data/contrastVOR.Rdata')
```

```{r data_prep_univariate}
# Each variable by treatment over study timeframe 
veg_univ <-
  veg_ra %>%
    select(block, pasture, patch, subpatch, transect, 
           TRT, year, Bare:IntWoody) %>%
    pivot_longer(cols = c(Bare:IntWoody), 
                 names_to = "spp", 
                 values_to = "cover") %>%
    group_by(TRT, pasture, year, spp) %>%
      summarize(Mean = mean(cover), 
                .groups = 'drop') %>%
    group_by(TRT, year, spp) %>%
    summarize(CoverMean = mean(Mean), 
              CoverSE = sd(Mean)/sqrt(n()), 
              .groups = 'drop') %>%
    rename("Management" = "TRT") %>%
  # filter(! spp %in% c('IntC4', 'IntWoody')) %>%
    mutate(group = case_when(
      spp %in% c('Bare', 'Dead', 'Robel', 'LitterDepth') ~ "Structural", 
      spp %in% c('POPR', 'BRIN') ~ "Focal", 
      TRUE ~ "Functional" )) %>%
    mutate(status = case_when(
      substr(spp, 1,3) == "Nat" ~ "Native",
      substr(spp, 1,3) == "Int" ~ "Exotic",
      TRUE ~ 'NA' )) %>%
  mutate(spp = recode(spp,
    "Bare" = "Bare ground", 
    'BRIN' = "Bromus inermis", 
    'Dead' = "Litter cover", 
    'IntC3' = 'Cool-season\ngrasses', 
    'IntForb' = 'Forbs', 
    'Intleg' ="Legumes", 
    'LitterDepth' = 'Litter depth', 
    'NatC3' = 'Cool-season\ngrasses', 
    'IntC4' = 'Warm-season\ngrasses', 
    'NatC4' = 'Warm-season\ngrasses', 
    'NatForb' = 'Forbs', 
    'Natleg' = 'Legumes', 
    'NatWoody' = "Woody\nplants", 
    'IntWoody' = "Woody\nplants", 
    'POPR' = "Poa pratensis", 
    'Robel' = 'Robel height'  ))
```

# Key take-aways 

* Initial composition across treatment groups was not consistent. 
The Refuge pastures that were assigned to continuous grazing without fire had substantially more smooth brome and fewer native forbs and legumes. 

* Few plant functional groups showed trends that can reasonably be attributed to management. 
These are limited to increased native forbs and cool-season grasses on PBG pastures.
Only PBG pastures showed consistent trends over time in terms of overall community composition. 

* Overall plant functional group diversity (beta diversity w/in treatment groups) was unaffected by management, which was unexpected. 

* Several structural variables followed expectations. 
Spatially-patchy fire increased both bare ground and litter on PBG pastures. 
Continuous grazing without fire accummulated litter. 

* Initially, spatial heterogeneity in vegetation structure was consistent across all pastures (low). 
Over time, conducting 40-ac patch burns increased patch contrast, as did rotational grazing. 


# Plant functional group composition & vegetation structure 

## Univariate trends--year by year

Tracking each response variable by treatment over the time period of the study. 

### Smooth brome and Kentucky bluegrass

```{r CSG_gg, fig.width=8, fig.height=3.5}
# POPR & BRIN  
  veg_univ %>%
    filter(group == 'Focal') %>%
    ggplot(aes(x = year)) + theme_bw(16) +
      geom_line(aes(y = CoverMean, 
                    color = Management), 
                position = position_dodge(width = 0.3)) +
      geom_errorbar(aes(ymin = CoverMean - CoverSE, 
                        ymax = CoverMean + CoverSE, 
                        color = Management), 
                    width = 0.25,
                    position = position_dodge(width = 0.3)) +
      geom_point(aes(y = CoverMean, 
                     fill = Management, 
                     shape = Management),
                 position = position_dodge(width = 0.3)) +
    labs(y = "Relative abundance\n(mean % ± s.e.)") +
    scale_shape_manual(values = c(21:24)) +
    scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
      facet_wrap(~spp, scales = "free_y") + 
    theme(strip.text = element_text(face = "italic"), 
          panel.grid.minor.x = element_blank())
```

Smooth brome was obviously disproportionately abundant on the Refuge pastures, 2-3x more abundant than other pastures. 
No evidence of change across any management type through study period. 
Kentucky bluegrass was generally about the same across all pastures and showed little change over the course of the study period. 

### Plant functional groups 

```{r funct_gg, fig.width=10}
# Functional groups
  veg_univ %>%
    filter(group == 'Functional') %>%
    mutate(spp = factor(spp, levels = c("Forbs", "Legumes", 
                                        "Cool-season\ngrasses", 
                                        "Warm-season\ngrasses", 
                                        "Woody\nplants"))) %>% 
    ggplot(aes(x = year)) + theme_bw(18) +
    geom_line(aes(y = CoverMean, 
                  color = Management), 
              position = position_dodge(width = 0.3)) +
    geom_errorbar(aes(ymin = CoverMean - CoverSE, 
                      ymax = CoverMean + CoverSE, 
                      color = Management), 
                  width = 0.25,
                  position = position_dodge(width = 0.3)) +
    geom_point(aes(y = CoverMean, 
                   fill = Management, 
                   shape = Management),
               position = position_dodge(width = 0.3)) +
    labs(y = "Relative abundance\n(mean % ± s.e.)") +
    scale_shape_manual(values = c(21:24)) +
    scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
    facet_grid(status~spp, scales = "free_y") +
    theme(axis.text.x = element_text(angle = 33, hjust = 1), 
          panel.grid.minor.x = element_blank())
```

Native forbs and legumes were markedly lower in the Refuge than other management types, which were not terribly different from each other outside of the Refuge. 
The Wagon Wheel was consistently higher in native woody plants. 

Few functional groups changed over the course of the study period. 
2019-2020 changes seem isolated and are likely attributable to precipitation (overall lower productivity in 2020). 

The increase in native cool-season grasses in PBG pastures is consistent with, and reflected in, the ordination as well. 
This is probably a real treatment effect. 

### Vegetation structure

```{r struct_gg, fig.width=8}
# Vegetation structure
  veg_univ %>%
    filter(group == 'Structural') %>%
    mutate(spp = case_when(
      spp %in% c("Bare ground", "Litter cover") ~ paste(spp, "(%)"), 
      spp == 'Litter depth' ~ paste(spp, "(mm)"), 
      spp == 'Robel height' ~ paste(spp, "(dm)")
    )) %>%
    ggplot(aes(x = year)) + theme_bw(16) +
    geom_line(aes(y = CoverMean, 
                  color = Management), 
              position = position_dodge(width = 0.3)) +
    geom_errorbar(aes(ymin = CoverMean - CoverSE, 
                      ymax = CoverMean + CoverSE, 
                      color = Management), 
                  width = 0.25,
                  position = position_dodge(width = 0.3)) +
    geom_point(aes(y = CoverMean, 
                   fill = Management, 
                   shape = Management),
               position = position_dodge(width = 0.3)) +
  labs(y = "Mean ± s.e.") +
    scale_shape_manual(values = c(21:24)) +
    scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
    facet_wrap(~spp, scales = "free_y") +
  theme( panel.grid.minor.x = element_blank())
```

Increases in both litter and bare ground on PBG pastures is only possible with patchy fire, and can be interpreted as a treatment effect. 
Likewise, increasing litter cover & depth on the Refuge is an expected result, and probably for rotational pastures, as well. 

## Changes over study period 

Comparing final to initial values for each measured variable.
Not sure how insightful these data are, given the influence of the 2020 drought. 
Hard to attribute any reductions, especially, to management. 

```{r}
# Calculating changes
veg_ch <-
  veg_ra  %>%   
    mutate(bookend = case_when(
      TRT == "Rotational" & year == '2018' ~ "Initial", 
      TRT != "Rotational" & year == '2017' ~ "Initial", 
      year == '2020' ~ "Final", 
      TRUE ~ 'NA'      )) %>% 
    filter(bookend %in% c("Initial", "Final")) %>%
  select(block, pasture, transect, 
         TRT, bookend, Bare:IntWoody) %>%
  pivot_longer(cols = c(Bare:IntWoody), 
               names_to = "var", 
               values_to = "value") %>%
  pivot_wider(names_from = bookend, 
              values_from = value) %>%
    mutate(diff = Final - Initial) %>%
  group_by(TRT, pasture, var) %>%
  summarize(change = mean(diff), 
            .groups = 'drop') %>%
  group_by(TRT, var) %>%
  summarize(ChangeMean = mean(change, na.rm = T), 
            ChangeSE = sd(change, na.rm = T)/sqrt(n()), 
            .groups = 'drop') %>%
  rename("Management" = "TRT") %>% 
    mutate(group = case_when(
      var %in% c('Bare', 'Dead', 'Robel', 'LitterDepth') ~ "Structural", 
      var %in% c('POPR', 'BRIN') ~ "Focal", 
      TRUE ~ "Functional" )) %>%
    mutate(status = case_when(
      substr(var, 1,3) == "Nat" ~ "Native",
      substr(var, 1,3) == "Int" ~ "Exotic",
      TRUE ~ 'NA' )) %>%
    mutate(var = recode(var,
                        "Bare" = "Bare ground", 
                        'BRIN' = "Bromus inermis", 
                        'Dead' = "Litter cover", 
                        'IntC3' = 'Cool-season\ngrasses', 
                        'IntForb' = 'Forbs', 
                        'Intleg' ="Legumes", 
                        'LitterDepth' = 'Litter depth', 
                        'NatC3' = 'Cool-season\ngrasses', 
                        'IntC4' = 'Warm-season\ngrasses', 
                        'NatC4' = 'Warm-season\ngrasses', 
                        'NatForb' = 'Forbs', 
                        'Natleg' = 'Legumes', 
                        'NatWoody' = "Woody\nplants", 
                        'IntWoody' = "Woody\nplants", 
                        'POPR' = "Poa pratensis", 
                        'Robel' = 'Robel height'  ))
```

### Smooth brome and Kentucky bluegrass 

```{r focal_ch_gg, fig.height = 4}
# POPR & BRIN  
veg_ch %>%
  filter(group == 'Focal') %>%
  ggplot(aes(x = var)) + theme_bw(16) +
  geom_hline(yintercept = 0, lty = 3, col = "black") + 
  geom_errorbar(aes(ymin = ChangeMean - ChangeSE, 
                    ymax = ChangeMean + ChangeSE, 
                    color = Management), 
                width = 0.25,
                position = position_dodge(width = 0.3)) +
  geom_point(aes(y = ChangeMean, 
                 fill = Management, 
                 shape = Management),
             position = position_dodge(width = 0.3)) +
  scale_shape_manual(values = c(21:24)) +
    scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  labs(x = "Species", 
       y = "Change over study\n(mean % ± s.e.)") + 
  theme(axis.text.x = element_text(face = "italic"))
``` 

### Plant functional groups

```{r funct_ch_gg}
# FUnctional groups
veg_ch %>%
  filter(group == 'Functional') %>%
  mutate(spp = factor(var, levels = c("Forbs", "Legumes", 
                                      "Cool-season\ngrasses", 
                                      "Warm-season\ngrasses", 
                                      "Woody\nplants"))) %>% 
  ggplot(aes(x = var)) + theme_bw(14) +
  geom_hline(yintercept = 0, lty = 3, col = "black") + 
  geom_errorbar(aes(ymin = ChangeMean - ChangeSE, 
                    ymax = ChangeMean + ChangeSE, 
                    color = Management), 
                width = 0.25,
                position = position_dodge(width = 0.3)) +
  geom_point(aes(y = ChangeMean, 
                 fill = Management, 
                 shape = Management),
             position = position_dodge(width = 0.3)) +
  scale_shape_manual(values = c(21:24)) +
    scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  labs(x = "Plant functional group", 
       y = "Change over study\n(mean % ± s.e.)") +
  facet_wrap(~status, scales = "free_y", ncol = 1) 
```

### Vegetation structure 

```{r struct_ch_gg, fig.height = 4} 
# Vegetation structure
veg_ch %>%
  filter(group == 'Structural') %>%
  ggplot(aes(x = var)) + theme_bw(16) +
  geom_hline(yintercept = 0, lty = 3, col = "black") + 
  geom_errorbar(aes(ymin = ChangeMean - ChangeSE, 
                    ymax = ChangeMean + ChangeSE, 
                    color = Management), 
                width = 0.25,
                position = position_dodge(width = 0.3)) +
  geom_point(aes(y = ChangeMean, 
                 fill = Management, 
                 shape = Management),
             position = position_dodge(width = 0.3)) +
  scale_shape_manual(values = c(21:24)) +
    scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  labs(x = "Structural component", 
       y = "Change over study\n(mean ± s.e.)")
```

## Multivariate analyses

### Community dissimilarity

```{r mv_data_rep, message }
# Multivariate analysis
  # Focus on first and final years (2018 & 2020)
    veg_d <- 
      veg_all %>%
      mutate(bookend = case_when(
              TRT == "Rotational" & year == '2018' ~ "Initial", 
              TRT != "Rotational" & year == '2017' ~ "Initial", 
              year == '2020' ~ "Final", 
              TRUE ~ 'NA'      )) %>% 
        filter(bookend %in% c("Initial", "Final")) %>%
      mutate(YrTrt = paste0(bookend,"_", TRT))
  # spp data only
    comm_d <- 
      veg_d %>%
      select(NatC3:IntWoody)  %>%
      select(-IntWoody, -IntC4) %>%
      replace(is.na(.), 0) 
  # Calculate distance matrix
    veg_m <- vegdist(comm_d, "bray")
 # perform ordination
    # Fit 
   veg_ord <- suppressMessages(
                capscale(comm_d ~ 1, distance = 'bray', 
                         metaMDSdist = TRUE, trace=FALSE)
                  )
   # Test groups
     veg_fit <- envfit(veg_ord ~ YrTrt + Bare + Dead + Robel + 
                         LitterDepth + POPR + BRIN, 
                       veg_d, strata = veg_d$block, 
                       choices = c(1:3), 199)
     pw <- pairwise.factorfit(veg_ord, veg_d$YrTrt, 
                              perm = 199, p.method = "holm" )$p.value %>%
            round(., 2)
     bd <- betadisper(veg_m, veg_d$YrTrt, bias.adjust=TRUE  )

  # Getting scores 
  veg_gg <- gg_ordiplot(veg_ord, groups = veg_d$YrTrt, 
                         plot=FALSE)
  veg_scores <- lst(species=scores(veg_ord, display = "species", 
                                   choices = c(1:2)) %>%
                            as.data.frame %>%
                            as_tibble(rownames="group"), 
                    sites=scores(veg_ord, display = "sites", 
                                 choices = c(1:2)) %>%
                      as_tibble() %>%
                      bind_cols(YrTrt = veg_d$YrTrt) %>%
                      separate(YrTrt, into=c('bookend', 'Trt'), 
                               sep = "_", remove = F) %>%
                      mutate(bookend = factor(bookend, 
                                              levels = c("Initial", "Final"))), 
                    spiders = veg_gg$df_spiders %>% 
                           rename(MDS1 = x, MDS2 = y) %>%
                           as_tibble %>%
                           separate(Group, into=c('bookend', 'Trt'), 
                                    sep = "_", remove = F) %>%
                      mutate(bookend = factor(bookend, 
                                              levels = c("Initial", "Final"))), 
                    vectors = scores(veg_fit,  "vectors") %>%
                              as.data.frame %>% 
                              round(3) %>%
                              as_tibble(rownames="gradient"), 
                    groups = pw, 
                    beta = bd$distances)
```

***Ordination results.--***
Very slight changes are apparent among groups and over time in ordination space. 
Continuously-grazed pastures appear the most widely distributed at the beginning of the study, with several pastures associated with Native Woody Plants; the group contracts in the final period of the study. 
Unburned treatments (rotationally-grazed and continuously-grazed pastures) converge in ordination space, while both PBG treatments appear similar to each other in both timesteps and appear to move along a gradient towards grater native plant groups. 

```{r ord_gg, fig.height= 4, fig.width=8}  
# Plot ordination 
  ord_gg <- ggplot() + theme_bw(16) + 
    labs(x="MDS Axis 1", 
         y="MDS Axis 2") + 
    geom_vline(xintercept = 0, lty=3, color="darkgrey") +
    geom_hline(yintercept = 0, lty=3, color="darkgrey") +
    theme(panel.grid=element_blank())
 ord_gg + 
    geom_segment(data=veg_scores$spiders, 
                 aes(x=cntr.x, y=cntr.y,
                     xend=MDS1, yend=MDS2, 
                     color=Trt), 
                 alpha = 0.5,
                 size=1.2) + 
    geom_point(data=veg_scores$spiders, 
               aes(x=MDS1, y=MDS2,
                   fill = Trt, 
                   shape = Trt), 
               size=2) +
    geom_label(data=veg_scores$spiders ,
               aes(x=cntr.x, y=cntr.y, 
                   label=Trt, 
                   color=Trt), 
               fontface="bold", size=4,
               label.size = 0, 
               label.r = unit(0.5, "lines")) +
    geom_label(data=veg_scores$species %>%
                 filter(! group %in% c("IntWoody", "IntC4")), 
               aes(x=MDS1*0.225, 
                   y=MDS2*0.225, 
                   label=group), 
               label.padding=unit(0.1,"lines"),
               label.size = 0, 
               fontface="bold", color="darkred") +
    scale_shape_manual(values = c(21:24)) +
     scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
facet_wrap(~bookend) + 
    theme(legend.position = 'none')
```

***Pairwise group comparisons.--***
Initially, most treatments were different from each other on account of the treatments not being randomly assigned across blocks. 
Only plant functional group composition in pastures in the two patch-burn grazing treatments were similar (Initial 2 burns/yr vs. final 2 burns/yr, P = `r pw[5,5]`). 
Continuously-grazed pastures were different from those in the rotational grazing treatment (P = `r pw[7,7]`) and both PBG treatments (1 burn/yr, P = `r pw[6,5]`; 2 burns/yr, P = `r pw[6,6]`). 

By the final year, plant functional group composition in PBG treatments had diverged (1 burn/yr vs. 2 burns/yr, P = `r pw[1,1]`). 
Composition in both PBG treatments changed over the course of the study (1 burn/yr, P = `r pw[4,1]`; 2 burns/yr, P = `r pw[5,2]`). 

But composition of continuously-grazed pastures was no different than when the study started (P = `r pw[6,3]`) and was not different from the final composition of rotationally-grazed pastures, either (P = `r pw[3,3]`). 
Over the course of the study, composition of continuously-grazed pastures remained different than both PBG treatments (1 burn/yr, P = `r pw[2,1]`; 2 burns/yr, P = `r pw[2,2]`).

Composition of rotationally-grazed pastures did shift over the course of the study (P = `r pw[7,4]`) and was different than both PBG treatments (1 burn/yr, P = `r pw[3,1]`; 2 burns/yr, P = `r pw[3,2]`).

### Trajectory analysis 

Arrows show how pasture centroids moved in ordination space from initial to final. 
Only PBG pastures moved consistently within treatment groups; this movement was away from introduced plant groups and towards native groups. 
General trend for Refuge pastures to increase towards native groups, as well, but to a lesser extent, consistent with univariate trends. 

```{r trajectory}
# Trajectory analysis  
veg_scores$sites %>%
  bind_cols(pasture = veg_d$pasture, .) %>%
  pivot_longer(cols = c(MDS1, MDS2), 
               names_to = 'axis', 
               values_to = 'score') %>%
  group_by(Trt, pasture, bookend, axis) %>%
    summarize(centroid = mean(score), 
              error = sd(score)/sqrt(n()),
              .groups = 'drop') %>%
  select(-error) %>%
  pivot_wider(names_from = c(bookend, axis), 
              values_from = centroid) %>%
  ggplot() + theme_bw(16) + 
  geom_segment(aes(x=Initial_MDS1, y=Initial_MDS2,
                   xend=Final_MDS1, yend=Final_MDS2, 
                   color=Trt), 
               arrow = arrow(length = unit(0.03, "npc")),
                 size=1.2)  +
  geom_label(data=veg_scores$species %>%
               filter(! group %in% c("IntWoody", "IntC4")), 
             aes(x=MDS1*0.5, 
                 y=MDS2*0.5, 
                 label=group), 
             label.padding=unit(0.1,"lines"),
             label.size = 0, 
             fontface="bold", color="darkred") +
  coord_cartesian(xlim = c(-0.75, 2))  +   
    labs(x="MDS Axis 1", 
         y="MDS Axis 2") + 
    geom_vline(xintercept = 0, lty=3, color="darkgrey") +
    geom_hline(yintercept = 0, lty=3, color="darkgrey") +
    theme(panel.grid=element_blank()) +
    scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) 
```
  
### Beta diversity 

Beta diversity is a measure of the area a group occupies in ordination space and can be interpreted as vegetation heterogeneity within groups. 
Management practices meant to increase spatial heterogeneity should increase beta diversity, while practices meant to reduce spatial heterogeneity should also show lower beta diversity (tigher groups in ordination space). 

Little effect was apparent over the course of the CGREC trials. 
Pastures at the Refuge started at relatively high spatial heterogeneity, which fell in line with other pastures following four years of continuous grazing. 

```{r beta_gg}
# Plotting changes in beta dispersion
  bind_cols(pasture = veg_d$pasture,
            YrTrt = veg_d$YrTrt, 
            disp = veg_scores$beta) %>% 
    separate(YrTrt, into = c("bookend", "Management"), sep = '_') %>%
    group_by(Management, bookend) %>%
    summarize(DispMean = mean(disp), 
              DispSE = sd(disp)/sqrt(n()), 
              .groups = 'drop') %>%
    mutate(bookend = factor(bookend, levels = c("Initial", "Final"))) %>%
    ggplot(aes(x = bookend)) + theme_bw(16) + 
    geom_line(aes(y = DispMean, 
                  color = Management, 
                  group = Management), 
              position = position_dodge(width = 0.1)) +
    geom_errorbar(aes(ymin = DispMean - DispSE, 
                      ymax = DispMean + DispSE, 
                      color = Management), 
                  width = 0.25, 
                  size = 1.1,
                  position = position_dodge(width = 0.1)) +
    geom_point(aes(y = DispMean, 
                   shape = Management, 
                   fill = Management), 
               size = 3, stroke = 1.25,
               position = position_dodge(width = 0.1)) + 
    coord_cartesian(xlim = c(1.4,1.6)) + 
    labs(x = "Data period", 
         y = "Within-group diversity") +
    scale_shape_manual(values = c(21:24)) +
    scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) 
```

# Spatial heterogeneity 

In this instance, *spatial heterogeneity* is defined as **patch contrast**--*the degree of difference among patches for several vegetation structure variables*. 
Patch contrast is measured as the amount of variance attributable to the patch term in a random-effect linear regression model fit for each response variable. 

Spatial heterogeneity tended to be greatest among PBG pastures with a single, 40-ac burn each spring. 
Lowest heterogeneity tended to be among PBG pastures in which two burns were planned each year, but summer fires often did not occur. 
Continuously-grazed pastures without fire tended to have low heterogeneity, while rotationally-grazed pastures generally increased in heterogeneity since the initiation of the treatment.

```{r contrast_calc, eval=FALSE}
veg_long <- 
  veg_all %>%
    pivot_longer(c(Bare:LitterDepth), 
                 names_to = 'response', 
                 values_to = 'values' ) 

# Check distributions
  veg_long %>% 
  ggplot(aes(x=log(values+1))) + theme_bw(16) + 
    geom_histogram(aes(y=..density..),      
                   binwidth=0.1,
                   colour="black",
                   fill="lightgreen") + 
    facet_wrap(~response, scales = 'free')

# Patch contrast
library(foreach)
  foreach(r=1:length(unique(veg_long$response))) %:%
  foreach(y=1:length(unique(veg_long$year))) %:%
  foreach(p=1:length(unique(veg_long$pasture))) %do% {
    d = filter(veg_long, 
               response == unique(veg_long$response)[r],
               year == unique(veg_long$year)[y], 
               pasture == unique(veg_long$pasture)[p])
    lme4::lmer(log(values + 1) ~ 
                 (1|patch), 
             data= d,
             REML = FALSE) %>% 
      lme4::VarCorr(. ) %>%
      as_tibble() %>%
      mutate(response = unique(d$response),
             treatment = unique(d$TRT), 
             pasture = unique(d$pasture), 
             year = unique(d$year)) %>%
      rename(term = grp, 
             variance = sdcor) %>%
      select(response, treatment, pasture, year, term, variance) %>%
      bind_rows(contrastVOR) -> contrastVOR
  }

# save(contrastVOR, file = './data/contrastVOR.Rdata')
```

```{r contrast_gg, fig.width=8}
contrastVOR %>%
  pivot_wider(names_from = term, 
              values_from = variance) %>%
  group_by(response, treatment, year) %>%
  summarize(PatchMean = mean(patch), 
            PatchSE = sd(patch)/sqrt(n()), 
            .groups = 'drop') %>%
  mutate(response = recode(response, 
                           Bare = 'Bare ground cover', 
                           Dead = 'Litter cover', 
                           LitterDepth = 'Litter depth', 
                           Robel = 'Visual Obstruction')) %>% 
  rename(Management = treatment) %>%
  ggplot(aes(x = year, 
             color = Management)) + theme_bw(14) + 
    geom_path(aes(y = PatchMean), 
              position = position_dodge(width = 0.1) ) + 
    geom_errorbar(aes(ymin = PatchMean - PatchSE, 
                      ymax = PatchMean + PatchSE), 
                  width = 0.2, 
                  position = position_dodge(width = 0.1)) + 
    geom_point(aes(y = PatchMean, 
                   fill = Management,
                   shape = Management), 
               size = 3, 
               color = "black", 
              position = position_dodge(width = 0.1) ) +
  labs(x = 'Year', 
       y = "Spatial heterogeneity") + 
  scale_shape_manual(values = c(21:24)) + 
  scale_color_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
  scale_fill_manual(values = wesanderson::wes_palette("Zissou1")[c(1:3,5)]) + 
    facet_wrap(~response, scales = "free_y") +
  theme(panel.grid.minor.x = element_blank())
```

